Appendix A : Dipole-dipole interaction between two spins in solid

### 2022-10-03 Manuscrit Thèse/chapter 2/Chapter 2.tex
ghuk
bikldf

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/tronche_de_spin}
\caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
\label{repr NV P1 VH}
\end{figure}

@@

The second reason comes from the nature of the sample used. We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers. We found that the dip at 55 G was present on all CVD samples ($\sim$ 5 samples) used, and none of the HPHT samples ($\sim$ 20 samples). The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below. We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.

@@

For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ one.

@@

This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the  .

@@

This procedure will be discussed in more details in the next chapter. We find here that the linewidth of VH$^-$ and War1 are respectively [REF] and [REF] MHz.

@@

Finally, the third tentative estimate we can make is that of the dark spins concentration. We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV]. We can then, knowing the NV concentration, try to estimate the concentration of the dark spins. There are several caveats though:

@@

With all this considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):

For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.

@@

We can see that the VH$^-$ line in particular completely disappear on sample CVD-B, which we attribute to the additional annealing. This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.

@@

An unexpected side-study was done on HPHT samples. Indeed, the CVD samples used previously are grown by homoepithaxy on top of a diamond substrate. In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer. We Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.

@@

To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations. This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up. It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.

We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy than the [111] scans for weak magnetic field.

Finally we have seen an application of the CR to detect optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.

There still remains a lot of open questions and there are several experiments that could be build upon these results. I will list some of them here.

@@

We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$. This kind of values are not too far from ODMR contrast with NV ensembles, which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin. The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center. Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.

There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when you want to polarize or readout the VH$^-$ state, and out of resonance when you want to coherently manipulate the VH$^-$. This means that the magnetic field needs to be switched faster than the spin dynamics you want to probe. This might be doable for the population dynamics $T_1$ but this is probably too slow to probe the $T_2$ time.

The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance. The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$) is out of resonance. The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition. While the polarization on this transition, and the back-action it would produce on the NV centers, is not as strong as if you were manipulating the $\ket{0} \to \ket{+1}$ transition, it should still produce a measurable signal.

@@

We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$. This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope. We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing. We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$. It is however unclear to me whether this detection scheme would be sensitive enough to detect single dark spins.

>>>>

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{Figures/tronche_de_spin}
\caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
\label{repr NV P1 VH}
\end{figure}

@@

The second reason comes from the nature of the sample used. We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers. We found that the dip at 55 G was present on all CVD samples ($\sim$ 5 samples) used, and none of the HPHT samples ($\sim$ 20 samples). The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below. We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the growth process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.

@@

For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively.

@@

This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the dark spins Hamiltonian parameters, linewidth and concentration.

@@

This procedure will be discussed in more details in the next chapter. We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in [REF]. The linewidth $W=2.7$ MHz (half witdth at half maximum) of the NV centers was measured from ODMR (see Fig. [REF]). It includes all three hyper-fine transistions}.

@@

Finally, the third estimate we can make is that of the dark spins concentration. We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV]. We can then, knowing the NV concentration, try to estimate the concentration of the dark spins. There are however several caveats:

@@

From these considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):

For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose a dark to bright NV ratio of 0.25 as found in \citep{choi2017depolarization}.

@@

We can see that the VH$^-$ line in particular completely disappears on sample CVD-B, which we attribute to the additional annealing. This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.

@@

An unexpected side-study was done on HPHT samples. Indeed, the CVD samples used previously are grown by homoepithaxy on top of a diamond substrate. In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer. We will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.

@@

To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations. This detection can be significantly more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up. This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would probably be to characterize the environment of samples with high NV density.

We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan which provides a significantly less noisy signal than the [111] scans for weak magnetic field.

Finally we have seen an application of the CR to detect and characterize optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.

There still remains a lot of open questions and there are several experiments that could be build upon these results. Some of these questions are listed below.

@@

We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$. This value is within one order of magnitude from ODMR contrast with NV ensembles (typically $<5 \%$), which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin. The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center. Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.

There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when initializing or reading out the VH$^-$ state, and out of resonance to coherently manipulate the VH$^-$. This means that the magnetic field needs to be switched faster than the probed spin dynamics, which may be doable for the population dynamics $T_1$ but is likely too slow to probe the $T_2$ time.

The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance. The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$ and $\ket{0} \to \ket{+1}$ for the NV) is out of resonance. The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition. Further studies are needed to assess the effectiveness of this scheme.

@@

We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$. This concentration is about two orders of magnitude too high to contemplate isolating a single emitter with a diffraction-limited microscope. We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing. We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$. It is however still unclear whether the present detection scheme is sensitive enough to detect single VH$^-$.
